---
title: "Numerical Simulations"
author: "Tianyi Liu"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
pacman::p_load(tidyverse, stringr, purrr, furrr, magrittr, zoo, patchwork)
source("./core_algorithm_functions_test.R")
source("./matrix_initiation_functions.R")
source("./utility_functions.R")
```

## Prelim: What is The Expected Norm Under Random Guesses?
```{r random.guess}
set.seed(1000)
## we have a 5 * 50 C matrix; want to see what is the exprected difference under random guesses
## if both truth and guesses are generated from a uniform distribution
n_subjects = 1000
n_cell_types = 5

unif_vec <- vector()
unif_cors <- vector()
for (i in 1:1000) {
  a_array = runif(n_subjects * n_cell_types, 0, 1)
  b_array= runif(n_subjects * n_cell_types, 0, 1)
  a = matrix(a_array, nrow=n_cell_types, ncol=n_subjects)
  b = matrix(b_array, nrow=n_cell_types, ncol=n_subjects)
  a = apply(a, 2, function(x) x/sum(x))
  b = apply(b, 2, function(x) x/sum(x))
  unif_vec[i] = norm(a - b, type="F")/sqrt(n_subjects)
  unif_cors[i] = colwise_cor(a, b, "pearson")
}

print(paste0("Uniform normalized difference norm: ", round(mean(unif_vec), 4)))
print(paste0("Uniform correlation: ", round(mean(unif_cors), 4)))

## from a dirichlet distribution
dirich_vec <- vector()
dirich_cors <- vector()
for (i in 1:1000) {
  c = matrix(t(rdirichlet(n_subjects, rep(1, n_cell_types))),
             n_cell_types,
             n_subjects,
             byrow=FALSE)
  d = matrix(t(rdirichlet(n_subjects, rep(1, n_cell_types))),
             n_cell_types,
             n_subjects,
             byrow=FALSE)
  dirich_vec[i] = norm(c - d, type="F")/sqrt(n_subjects)
  dirich_cors[i] = colwise_cor(c, d, "pearson")
}

print(paste0("Dirichlet normalized difference norm: ", round(mean(dirich_vec), 4)))
print(paste0("Dirichlet correlation: ", round(mean(dirich_cors), 4)))
```
Hence, we draw the true CT proportions from a Dirichlet distirbution to allow for more expected variation.

```{r}
set.seed(100)

# set the constants
n_genes <- 100
n_cell_types <- 5
n_subjects <- 1000
n_known_genes <- 60
n_good_cell_types <- 4

## generate a G
G_true <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)

## generate a C
C_true <- get_C(n_cell_types, n_subjects)

## generate an error matrix
E <- matrix(runif(n_genes * n_subjects), n_genes, n_subjects)

## generate an M
M <- G_true %*% C_true + E  

## get G0 from G
G0 <- matrix(0L, n_genes, n_cell_types)
G0[1:n_known_genes, 1:n_good_cell_types] <- G_true[1:n_known_genes, 1:n_good_cell_types]

## Get Delta; will be useful for getting all the norms
Delta <- get_Delta(n_cell_types, n_good_cell_types, n_genes, n_known_genes)
```

## Experiment 1:
We pull one value for each of the tuning parameter around the suitable range we discovered. Then, we contrast the solution of the two methods iteration by iteration, using `G_true` and `C_true` as initial values:

```{r exp1}
## one value for each parameter
# a = 2^16
# x = 16
# b = 2^20

a = 2^10
x = 2^10
b = 2^10

## run one numerical simulation
solve_direct <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_true,
  C_init = C_true,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'direct',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)

solve_auxiliary <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_true,
  C_init = C_true,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'auxiliary',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)
```

Collect the results:
```{r}
## output all the residuals we collected
direct_n_iter = solve_direct$n_iter 
direct_obj = solve_direct$obj
direct_M_res = solve_direct$M_res_vec
direct_CG_res = solve_direct$CG_res_vec 
direct_CG_secondary_res = solve_direct$CG_secondary_res_vec 
direct_C_cors = solve_direct$C_cors
direct_C_spears = solve_direct$C_spears

direct_df <- tibble(
  iter = 1:direct_n_iter,
  obj = direct_obj,
  M_res = direct_M_res,
  CG_res = direct_CG_res,
  CG_secondary_res = direct_CG_secondary_res,
  C_cors = direct_C_cors,
  C_spears = direct_C_spears
)

aux_n_iter = solve_auxiliary$n_iter 
aux_obj = solve_auxiliary$obj
aux_M_res = solve_auxiliary$M_res_vec
aux_CG_res = solve_auxiliary$CG_res_vec 
aux_CG_secondary_res = solve_auxiliary$CG_secondary_res_vec 
aux_C_cors = solve_auxiliary$C_cors
aux_C_spears = solve_auxiliary$C_spears

aux_df <- tibble(
  iter = 1:aux_n_iter,
  obj = aux_obj,
  M_res = aux_M_res,
  CG_res = aux_CG_res,
  CG_secondary_res = aux_CG_secondary_res,
  C_cors = aux_C_cors,
  C_spears = aux_C_spears
)
```

We plot all the results:
```{r}
p0 <- direct_df %>% 
  ggplot(aes(x=iter, y=obj)) + 
  geom_line(col="green")

p0 <- p0 + 
  geom_line(data=aux_df, aes(x=iter, y=obj), col="red") +
  labs(
    title="Objectives",
    x="Iteration",
    y="Objective Function"
  )

p1 <- direct_df %>% 
  ggplot(aes(x=iter, y=M_res)) + 
  geom_line(col="green")

p1 <- p1 + 
  geom_line(data=aux_df, aes(x=iter, y=M_res), col="red") +
  labs(
    title="Residual of M",
    x="Iteration",
    y="||M - M_hat||_F"
  )

# p2 <- direct_df %>% 
#   filter(row_number() %% 2 == 0) %>% 
#   ggplot(aes(x=iter, y=CG_res)) + 
#   geom_line(col="green")
# 
# p2 <- p2 + 
#   geom_line(data=aux_df %>% filter(row_number() %% 2 == 0), aes(x=iter, y=CG_res), col="red") +
#   labs(
#     title="Residual of G",
#     x="Iteration",
#     y="||G - G_hat||_F"
#   )

p2 <- direct_df %>%
  filter(row_number() %% 2 == 1) %>%
  ggplot(aes(x=iter, y=CG_res/sqrt(n_subjects))) +
  geom_line(col="green")

p2 <- p2 +
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_res/sqrt(n_subjects)), col="red") +
  labs(
    title="Residual of C",
    x="Iteration",
    y="||C - C_hat||_F"
  )

p3 <- direct_df %>%
  ggplot(aes(x=iter, y=C_spears)) +
  geom_line(col="green")

p3 <- p3 +
  geom_line(data=aux_df, aes(x=iter, y=C_spears), col="red") +
  labs(
    title="Mean Columnwise \n Spearman Correlations of C",
    x="Iteration",
    y="(1/n) * sum(cor(C, C_true))"
  )

p4 <- direct_df %>% 
  filter(row_number() %% 2 == 0) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p4 <- p4 + 
  geom_line(data=aux_df %>%  filter(row_number() %% 2 == 0), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of Fixed \nPart of G",
    x="Iteration",
    y="||Delta * G_hat - C_0||_F"
  )

p5 <- direct_df %>% 
  filter(row_number() %% 2 == 1) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p5 <- p5 + 
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of \nColumn Sum of C",
    x="Iteration",
    y="||1'C_hat - 1||_2"
  )

(p0 | p1 | p2)/(p3 | p4 | p5)
```

## Experiment 2:
We pull one value for each of the tuning parameter around the suitable range we discovered. Then, we contrast the solution of the two methods iteration by iteration, using two randomly generated `G_init` and `C_init` as initial values:

```{r exp2}
## change the initial values
## generate a G
G_init <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
C_init <- get_C(n_cell_types, n_subjects)

## one value for each parameter
a = 2^10
x = 2^10
b = 2^10

# a = 2^10
# x = 16
# b = 2^10

## run one numerical simulation
solve_direct <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_init,
  C_init = C_init,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'direct',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)

solve_auxiliary <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_init,
  C_init = C_init,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'auxiliary',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)
```

Collect the results:
```{r}
## output all the residuals we collected
direct_n_iter = solve_direct$n_iter 
direct_obj = solve_direct$obj
direct_M_res = solve_direct$M_res_vec
direct_CG_res = solve_direct$CG_res_vec 
direct_CG_secondary_res = solve_direct$CG_secondary_res_vec 
direct_C_cors = solve_direct$C_cors
direct_C_spears = solve_direct$C_spears

direct_df <- tibble(
  iter = 1:direct_n_iter,
  obj = direct_obj,
  M_res = direct_M_res,
  CG_res = direct_CG_res,
  CG_secondary_res = direct_CG_secondary_res,
  C_cors = direct_C_cors,
  C_spears = direct_C_spears
)

aux_n_iter = solve_auxiliary$n_iter 
aux_obj = solve_auxiliary$obj
aux_M_res = solve_auxiliary$M_res_vec
aux_CG_res = solve_auxiliary$CG_res_vec 
aux_CG_secondary_res = solve_auxiliary$CG_secondary_res_vec 
aux_C_cors = solve_auxiliary$C_cors
aux_C_spears = solve_auxiliary$C_spears

aux_df <- tibble(
  iter = 1:aux_n_iter,
  obj = aux_obj,
  M_res = aux_M_res,
  CG_res = aux_CG_res,
  CG_secondary_res = aux_CG_secondary_res,
  C_cors = aux_C_cors,
  C_spears = aux_C_spears
)
```

```{r}
p0 <- direct_df %>% 
  ggplot(aes(x=iter, y=obj)) + 
  geom_line(col="green")

p0 <- p0 + 
  geom_line(data=aux_df, aes(x=iter, y=obj), col="red") +
  labs(
    title="Objectives",
    x="Iteration",
    y="Objective Function"
  )

p1 <- direct_df %>% 
  ggplot(aes(x=iter, y=M_res)) + 
  geom_line(col="green")

p1 <- p1 + 
  geom_line(data=aux_df, aes(x=iter, y=M_res), col="red") +
  labs(
    title="Residual of M",
    x="Iteration",
    y="||M - M_hat||_F"
  )

# p2 <- direct_df %>% 
#   filter(row_number() %% 2 == 0) %>% 
#   ggplot(aes(x=iter, y=CG_res)) + 
#   geom_line(col="green")
# 
# p2 <- p2 + 
#   geom_line(data=aux_df %>% filter(row_number() %% 2 == 0), aes(x=iter, y=CG_res), col="red") +
#   labs(
#     title="Residual of G",
#     x="Iteration",
#     y="||G - G_hat||_F"
#   )

p2 <- direct_df %>%
  filter(row_number() %% 2 == 1) %>%
  ggplot(aes(x=iter, y=CG_res)) +
  geom_line(col="green")

p2 <- p2 +
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_res), col="red") +
  labs(
    title="Residual of C",
    x="Iteration",
    y="||C - C_hat||_F"
  )

p3 <- direct_df %>%
  ggplot(aes(x=iter, y=C_spears)) +
  geom_line(col="green")

p3 <- p3 +
  geom_line(data=aux_df, aes(x=iter, y=C_spears), col="red") +
  labs(
    title="Mean Columnwise \nSpearman Correlations of C",
    x="Iteration",
    y="(1/n) * sum(cor(C, C_true))"
  )

p4 <- direct_df %>% 
  filter(row_number() %% 2 == 0) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p4 <- p4 + 
  geom_line(data=aux_df %>%  filter(row_number() %% 2 == 0), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of Fixed \nPart of G",
    x="Iteration",
    y="||Delta * G_hat - C_0||_F"
  )

p5 <- direct_df %>% 
  filter(row_number() %% 2 == 1) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p5 <- p5 + 
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of \nColumn Sum of C",
    x="Iteration",
    y="||1'C_hat - 1||_2"
  )

(p0 | p1 | p2)/(p3 | p4 | p5)
```

### Experiment 3:
What if we make the penalty bigger?

```{r exp3}
## change the initial values
## generate a G
G_init <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
C_init <- get_C(n_cell_types, n_subjects)

## one value for each parameter
a = 2^20
x = 2^10
b = 2^20

# a = 2^10
# x = 16
# b = 2^10

## run one numerical simulation
solve_direct <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_init,
  C_init = C_init,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'direct',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)

solve_auxiliary <- PSMF_solve_test(
  M = M,
  G_0 = G0,
  G_init = G_true,
  C_init = C_true,
  n_markers = n_known_genes,
  n_good_cell_types = n_good_cell_types,
  alpha = a,
  xi = x,
  beta = b,
  eps = 1e-5,
  method = 'auxiliary',
  Delta = Delta,
  G_true = G_true,
  C_true = C_true
)
```

Collect the results:
```{r}
## output all the residuals we collected
direct_n_iter = solve_direct$n_iter 
direct_obj = solve_direct$obj
direct_M_res = solve_direct$M_res_vec
direct_CG_res = solve_direct$CG_res_vec 
direct_CG_secondary_res = solve_direct$CG_secondary_res_vec 
direct_C_cors = solve_direct$C_cors
direct_C_spears = solve_direct$C_spears

direct_df <- tibble(
  iter = 1:direct_n_iter,
  obj = direct_obj,
  M_res = direct_M_res,
  CG_res = direct_CG_res,
  CG_secondary_res = direct_CG_secondary_res,
  C_cors = direct_C_cors,
  C_spears = direct_C_spears
)

aux_n_iter = solve_auxiliary$n_iter 
aux_obj = solve_auxiliary$obj
aux_M_res = solve_auxiliary$M_res_vec
aux_CG_res = solve_auxiliary$CG_res_vec 
aux_CG_secondary_res = solve_auxiliary$CG_secondary_res_vec 
aux_C_cors = solve_auxiliary$C_cors
aux_C_spears = solve_auxiliary$C_spears

aux_df <- tibble(
  iter = 1:aux_n_iter,
  obj = aux_obj,
  M_res = aux_M_res,
  CG_res = aux_CG_res,
  CG_secondary_res = aux_CG_secondary_res,
  C_cors = aux_C_cors,
  C_spears = aux_C_spears
)
```

```{r}
p0 <- direct_df %>% 
  ggplot(aes(x=iter, y=obj)) + 
  geom_line(col="green")

p0 <- p0 + 
  geom_line(data=aux_df, aes(x=iter, y=obj), col="red") +
  labs(
    title="Objectives",
    x="Iteration",
    y="Objective Function"
  )

p1 <- direct_df %>% 
  ggplot(aes(x=iter, y=M_res)) + 
  geom_line(col="green")

p1 <- p1 + 
  geom_line(data=aux_df, aes(x=iter, y=M_res), col="red") +
  labs(
    title="Residual of M",
    x="Iteration",
    y="||M - M_hat||_F"
  )

# p2 <- direct_df %>% 
#   filter(row_number() %% 2 == 0) %>% 
#   ggplot(aes(x=iter, y=CG_res)) + 
#   geom_line(col="green")
# 
# p2 <- p2 + 
#   geom_line(data=aux_df %>% filter(row_number() %% 2 == 0), aes(x=iter, y=CG_res), col="red") +
#   labs(
#     title="Residual of G",
#     x="Iteration",
#     y="||G - G_hat||_F"
#   )

p2 <- direct_df %>%
  filter(row_number() %% 2 == 1) %>%
  ggplot(aes(x=iter, y=CG_res)) +
  geom_line(col="green")

p2 <- p2 +
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_res), col="red") +
  labs(
    title="Residual of C",
    x="Iteration",
    y="||C - C_hat||_F"
  )

p3 <- direct_df %>%
  ggplot(aes(x=iter, y=C_spears)) +
  geom_line(col="green")

p3 <- p3 +
  geom_line(data=aux_df, aes(x=iter, y=C_spears), col="red") +
  labs(
    title="Mean Columnwise \nSpearman Correlations of C",
    x="Iteration",
    y="(1/n) * sum(cor(C, C_true))"
  )

p4 <- direct_df %>% 
  filter(row_number() %% 2 == 0) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p4 <- p4 + 
  geom_line(data=aux_df %>%  filter(row_number() %% 2 == 0), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of Fixed \nPart of G",
    x="Iteration",
    y="||Delta * G_hat - C_0||_F"
  )

p5 <- direct_df %>% 
  filter(row_number() %% 2 == 1) %>% 
  ggplot(aes(x=iter, y=CG_secondary_res)) + 
  geom_line(col="green")

p5 <- p5 + 
  geom_line(data=aux_df %>% filter(row_number() %% 2 == 1), aes(x=iter, y=CG_secondary_res), col="red") +
  labs(
    title="Residual of \nColumn Sum of C",
    x="Iteration",
    y="||1'C_hat - 1||_2"
  )

(p0 | p1 | p2)/(p3 | p4 | p5)
```
