---
title: "First Simulations on New Algorithm"
author: "Tianyi Liu"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse)
source("./core_algorithm_functions.R")
source("./matrix_initiation_functions.R")
```

## Simulation Setups

 * Total number of genes: 20
 * Total number of cell types:5
 * Total number of subjects: 50
 * Total number of known genes: 12
 * Total number of good (known) cell types: 4
 
    * _Note: "good" means those cell types able to be sequenced_
    
 To simulate very nice errors, we just restrain each error term between $10^{-5}$ to $10^{-7}$ (basically no errors).

```{r}
## set a seed
set.seed(100)

## constants
n_genes <- 20
n_cell_types <- 5
n_subjects <- 50
n_known_genes <- 12
n_good_cell_types <- 4

## generate a G
G_true <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)

## generate a C
C_true <- get_C(n_cell_types, n_subjects)

## generate an error matrix
E <- matrix(runif(n_genes * n_subjects), n_genes, n_subjects)

## generate an M
M <- G_true %*% C_true + E  

## get G0 from G
G0 <- matrix(0L, n_genes, n_cell_types)
G0[1:n_known_genes, 1:n_good_cell_types] <- G_true[1:n_known_genes, 1:n_good_cell_types]
```

For the penalty, we will use values such that the result will be closed to what we wanted:

  * $\alpha = 200$
  * $\xi = 10$
  * $\beta = 200$
  
```{r}
alpha <- 200
xi <- 10
beta <- 200
```

### Method Comparison
To run the algorithm, we use both `direct` and `auxiliary` methods and compare their results. We begin with the `direct` method:
```{r}
G_init <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
C_init <- get_C(n_cell_types, n_subjects)

solve1 <- PSMF_solve(
  M=M, 
  G_0=G0, 
  G_init=G_init, 
  C_init=C_init,
  n_markers=n_known_genes, 
  n_good_cell_types=n_good_cell_types, 
  alpha=alpha, 
  xi=xi, 
  beta=beta,
  eps=1e-5, 
  method='direct')

solve2 <- PSMF_solve(
  M=M, 
  G_0=G0, 
  G_init=G_init, 
  C_init=C_init,
  n_markers=n_known_genes, 
  n_good_cell_types=n_good_cell_types, 
  alpha=alpha, 
  xi=xi, 
  beta=beta,
  eps=1e-5, 
  method='auxiliary')

norm(solve1$G_hat - solve2$G_hat, type="F")
norm(solve1$C_hat - solve2$C_hat, type="F")

norm(solve1$G_hat - G_true, type="F")
norm(solve2$G_hat - G_true, type="F")
norm(solve1$C_hat - C_true, type="F")
norm(solve2$C_hat - C_true, type="F")

plot(c(solve1$G_hat), c(solve2$G_hat))
plot(c(solve1$C_hat), c(solve2$C_hat))

plot(solve1$res_vec)
plot(solve2$res_vec)
```

Now we run 100 different initial values
```{r, message=FALSE, warning=FALSE}
n_sim <- 100

## initiate result data frame
results <- tibble(
  two_method_G_difference = rep(NaN, n_sim), 
  two_method_C_difference = rep(NaN, n_sim), 
  direct_G_residual = rep(NaN, n_sim), 
  auxiliary_G_residual = rep(NaN, n_sim), 
  direct_C_residual = rep(NaN, n_sim), 
  auxiliary_C_residual = rep(NaN, n_sim), 
)

## use the same Ms to compare two methods, but simulate a new M every single time
for (i in 1:n_sim) {
  
  G_true <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
  C_true <- get_C(n_cell_types, n_subjects) 
  E <- matrix(runif(n_genes * n_subjects), n_genes, n_subjects) 
  M <- G_true %*% C_true + E  
  G0 <- matrix(0L, n_genes, n_cell_types) 
  G0[1:n_known_genes, 1:n_good_cell_types] <- G_true[1:n_known_genes, 1:n_good_cell_types]
  
  G_init <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
  C_init <- get_C(n_cell_types, n_subjects)

  solve1 <- PSMF_solve(
    M=M, 
    G_0=G0, 
    G_init=G_init, 
    C_init=C_init,
    n_markers=n_known_genes, 
    n_good_cell_types=n_good_cell_types, 
    alpha=alpha, 
    xi=xi, 
    beta=beta,
    eps=1e-5, 
    method='direct')

  solve2 <- PSMF_solve(
    M=M, 
    G_0=G0, 
    G_init=G_init, 
    C_init=C_init,
    n_markers=n_known_genes, 
    n_good_cell_types=n_good_cell_types, 
    alpha=alpha, 
    xi=xi, 
    beta=beta,
    eps=1e-5, 
    method='auxiliary')
  
  results$two_method_G_difference[i] <- norm(solve1$G_hat - solve2$G_hat, type="F")
  results$two_method_C_difference[i] <- norm(solve1$C_hat - solve2$C_hat, type="F")

  results$direct_G_residual[i] <- norm(solve1$G_hat - G_true, type="F")
  results$auxiliary_G_residual[i] <- norm(solve2$G_hat - G_true, type="F")
  results$direct_C_residual[i] <- norm(solve1$C_hat - C_true, type="F")
  results$auxiliary_C_residual[i] <- norm(solve2$C_hat - C_true, type="F")
  
}

results %>% 
  ggplot(aes(x=two_method_G_difference)) + 
  geom_density(col="blue") +
  geom_rug()

results %>% 
  ggplot(aes(x=two_method_C_difference)) + 
  geom_density(col="blue") +
  geom_rug()

results %>% 
  ggplot(aes(x=direct_G_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results %>% 
  ggplot(aes(x=auxiliary_G_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results %>% 
  ggplot(aes(x=direct_C_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results %>% 
  ggplot(aes(x=auxiliary_C_residual)) + 
  geom_density(col="blue") +
  geom_rug()
```

### Initial Value Effects
The last thing we want to look at is how variable the results can be using different initial values:
```{r, message=FALSE, warning=FALSE}
n_initial_values <- 20

## initiate another result data frame
results2 <- tibble(
  two_method_G_difference = rep(NaN, n_initial_values), 
  two_method_C_difference = rep(NaN, n_initial_values), 
  direct_G_residual = rep(NaN, n_initial_values), 
  auxiliary_G_residual = rep(NaN, n_initial_values), 
  direct_C_residual = rep(NaN, n_initial_values), 
  auxiliary_C_residual = rep(NaN, n_initial_values), 
)

## use one true M
G_true <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
C_true <- get_C(n_cell_types, n_subjects) 
E <- matrix(runif(n_genes * n_subjects), n_genes, n_subjects) 
M <- G_true %*% C_true + E  

for (i in 1:n_initial_values) {

  G0 <- matrix(0L, n_genes, n_cell_types) 
  G0[1:n_known_genes, 1:n_good_cell_types] <- G_true[1:n_known_genes, 1:n_good_cell_types]
  
  G_init <- get_G(n_genes, n_subjects, n_cell_types, n_known_genes, n_good_cell_types)
  C_init <- get_C(n_cell_types, n_subjects)

  solve1 <- PSMF_solve(
    M=M, 
    G_0=G0, 
    G_init=G_init, 
    C_init=C_init,
    n_markers=n_known_genes, 
    n_good_cell_types=n_good_cell_types, 
    alpha=alpha, 
    xi=xi, 
    beta=beta,
    eps=1e-5, 
    method='direct')

  solve2 <- PSMF_solve(
    M=M, 
    G_0=G0, 
    G_init=G_init, 
    C_init=C_init,
    n_markers=n_known_genes, 
    n_good_cell_types=n_good_cell_types, 
    alpha=alpha, 
    xi=xi, 
    beta=beta,
    eps=1e-5, 
    method='auxiliary')
  
  results2$two_method_G_difference[i] <- norm(solve1$G_hat - solve2$G_hat, type="F")
  results2$two_method_C_difference[i] <- norm(solve1$C_hat - solve2$C_hat, type="F")

  results2$direct_G_residual[i] <- norm(solve1$G_hat - G_true, type="F")
  results2$auxiliary_G_residual[i] <- norm(solve2$G_hat - G_true, type="F")
  results2$direct_C_residual[i] <- norm(solve1$C_hat - C_true, type="F")
  results2$auxiliary_C_residual[i] <- norm(solve2$C_hat - C_true, type="F")
  
}

results2 %>% 
  ggplot(aes(x=two_method_G_difference)) + 
  geom_density(col="blue") +
  geom_rug()

results2 %>% 
  ggplot(aes(x=two_method_C_difference)) + 
  geom_density(col="blue") +
  geom_rug()

results2 %>% 
  ggplot(aes(x=direct_G_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results2 %>% 
  ggplot(aes(x=auxiliary_G_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results2 %>% 
  ggplot(aes(x=direct_C_residual)) + 
  geom_density(col="blue") +
  geom_rug()

results2 %>% 
  ggplot(aes(x=auxiliary_C_residual)) + 
  geom_density(col="blue") +
  geom_rug()
```

