---
title: "Analysis of Junjian's Data"
author: "Tianyi Liu"
date: "10/05/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir="/Users/gr8lawrence/Documents/Dissertation/conv_opt_code/cleaned_algorithm/PFSM/R")
pacman::p_load(tidyverse, stringr, purrr, furrr, magrittr, nnls)
source("./00_core_functions.R")
source("./01_matrix_initiation_functions.R")
source("./02_utility_functions.R")
```

## Seed Control
```{r set.seed}
set.seed(100)
```

## Read Data
We begin with reading in all the bulk, base, and proportion matrices:
```{r read.data, message=FALSE, warning=FALSE}
## directory names 
base_mtx_dir <- "./Real_data/Junjian_01/base_mtx"
bulk_dir <- "./Real_data/Junjian_01/bulk"
proportions_dir <- "./Real_data/Junjian_01/proportions"
output_dir <- "./Real_data/Junjian_01/output"

## all the base matrices
base_ABIS_logFC_a2s <- read_csv(paste(base_mtx_dir, "ABIS_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_immunoStates_logFC_a2s <- read_csv(paste(base_mtx_dir, "immunoStates_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_LM22_logFC_a2s <- read_csv(paste(base_mtx_dir, "LM22_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_tianyi_logFC_a2s <- read_csv(paste(base_mtx_dir, "tianyi_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)

## all the bulk matrices
bulk_markers_sd <- read_tsv(paste(bulk_dir, "bulk_markers_sd.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_markers_sd_q <- read_tsv(paste(bulk_dir, "bulk_markers_sd_q.txt", sep="/")) %>% rename(gene = GeneSymbol) 
bulk_nonmarkers_sd <- read_tsv(paste(bulk_dir, "bulk_nonmkrs_sd.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_nonmarkers_sd_q <- read_tsv(paste(bulk_dir, "bulk_nonmkrs_sd_q.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_Z0201 <- read_tsv(paste(bulk_dir, "Z0201_bulkCIBERSORT_a2s.txt", sep="/")) %>% rename(gene = GeneSymbol)

## all the proportion matrices
prop_tianyi <- read_csv(paste(proportions_dir, "CIBERSORTx_Job144_Results_tianyi.csv", sep="/")) 
prop_lm22 <- read_csv(paste(proportions_dir, "CIBERSORTx_Job145_Results_lm22.csv", sep="/"))
prop_immunoStates <- read_csv(paste(proportions_dir, "CIBERSORTx_Job146_Results_immuno.csv", sep="/"))
prop_abis <- read_csv(paste(proportions_dir, "CIBERSORTx_Job147_Results_abis.csv", sep="/"))
```

## Constants
We identify some constants from the data that are applciable to each of the matrices:
```{r get.constants}
n_cell_types <- 6
n_subjects <- 51
```

## Data Pre-processing
Next, we write a series of helper functions to obtain matrices for deconvolution:
```{r helper.function}
get_bulk <- function(bulk_df) {
  # subset the expression and get gene and subject names
  gene_names = bulk_df$gene
  bulk_exp_df = bulk_df %>% 
    select(2:(1 + n_subjects)) 
  subject_names = colnames(bulk_exp_df) %>% 
    substr(2, 5)
  
  # assemble the matrix
  mat = data.matrix(bulk_exp_df)
  rownames(mat) = gene_names
  colnames(mat) = subject_names
  return(mat)
}

#get_bulk(bulk_markers_sd)

get_base <- function(base_df) {
  gene_names = base_df$gene
  base_exp_df = base_df %>% 
    select(2:(1 + n_cell_types))
  cell_type_names = colnames(base_exp_df)
  
  mat = data.matrix(base_exp_df)
  rownames(mat) = gene_names
  colnames(mat) = cell_type_names
  return(mat)
}

#get_base(base_ABIS_logFC_a2s)

get_prop <- function(prop_df) {
  subject_names = prop_df$Mixture
  prop_real_df = prop_df %>% 
    select(2:(1 + n_cell_types))
  cell_type_names = colnames(prop_real_df)
  
  mat = data.matrix(prop_real_df) %>% t
  mat = apply(mat, 2, function(x) x/sum(x))
  rownames(mat) = cell_type_names
  colnames(mat) = subject_names
  return(mat)
}

#get_prop(prop_abis)
```

With the new functions we get all the 5 bulk matrices prepared:
```{r bulk.matrices}
bulk_markers_mat <- get_bulk(bulk_markers_sd)
bulk_markers_q_mat <- get_bulk(bulk_markers_sd_q) # what we might also want
bulk_nonmkrs_mat <- get_bulk(bulk_nonmarkers_sd)
bulk_nonmkrs_q_mat <- get_bulk(bulk_nonmarkers_sd_q) # what we want
bulk_Z0201_mat <- get_bulk(bulk_Z0201)
```

Then, we get the proportion:
```{r prop.matrices}
prop_abis_mat <- get_prop(prop_abis)
prop_lm22_mat <- get_prop(prop_lm22)
prop_immuno_mat <- get_prop(prop_immunoStates)
prop_tianyi_mat <- get_prop(prop_tianyi)

print(colSums(prop_tianyi_mat)[1:5])

# make a list of the four prop matrices
prop_ls <- list(prop_abis_mat, prop_lm22_mat, prop_immuno_mat, prop_tianyi_mat)
```

## Helper Function
Now we calculate $\hat{\mathbf{G}}_{\rm sub}$ from $\mathbf{M}_{\rm sub}$ (with only nonmarkers at average expression levels) and $\hat{\mathbf{C}}$ obtained from deconvolution and then calculate $\hat{\mathbf{M}}_{\rm sub} = \hat{\mathbf{G}}_{\rm sub}\hat{\mathbf{C}}$.

```{r helper.function2}
get_M_hat <- function(M_sub, C_hat) {
  t_G_sub_hat = matrix(0L, nrow=n_cell_types, ncol=nrow(M_sub)) 
  t_M_sub = t(M_sub)
  t_C_hat = t(C_hat)
  
  # calculate the NNLS solution for each column of C_hat
  for (j in 1:nrow(M_sub)) {
    nnls_sol = nnls(t_C_hat, t_M_sub[, j])
    t_G_sub_hat[, j] = nnls_sol$x
  }
  M_sub_hat = t(t_G_sub_hat) %*% C_hat 
  return(M_sub_hat)
}
```

```{r m.hat.list}
## get a list to store all four M hats from the non-marker genes with middle expression
M_hat_ls <- lapply(prop_ls, function(x) get_M_hat(bulk_nonmkrs_q_mat, x))
```

We want another function to process the weight $\mathbf{w} = (w_1, w_2, w_3, w_4)$ of all data. These four weights correspond to `abis`, `lm22`, `immunoStates`, `tianyi`. 

From SCDC, we wish to minimize $\left\| \mathbf{M} - \sum_{r = 1}^4 w_r\hat{\mathbf{M}}_r \right\|_{1, 1}$, with $w_1 + w_2 + w_3 + w_4 = 1$.

```{r ensemble1}
## helper function to process the errors
get_error <- function(M, M_hat_ls, wt_vec) {
  
  sum = matrix(0L, nrow(M), ncol(M))
  for (i in 1:length(M_hat_ls)) {
    sum = sum + wt_vec[i] * M_hat_ls[[i]]
  }
  
  return(sum(abs(M - sum)))
  
}

## test the function
# wt_vec <- rep(0.25, 4)
# M <- bulk_nonmkrs_q_mat
# print(get_error(M, M_hat_ls, wt_vec))
```


## Get The Best Combination of Weights
```{r ensemble2}
## the M matrix
M <- bulk_nonmkrs_q_mat

## set the grid
# w1_vec <- w2_vec <- w3_vec <- c(0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5)
w1_vec <- w2_vec <- w3_vec <- seq(0, 1, 0.05)

## set the slot for the best set of weights
wt_best = NULL
error_best = Inf

for (i in 1:length(w1_vec)) {
  for (j in 1:length(w2_vec)) {
    for (k in 1:length(w3_vec)) {
      
      w1 <- w1_vec[i]
      w2 <- w2_vec[j]
      w3 <- w3_vec[k]
      w4 <- 1 - (w1 + w2 + w3)
      
      if (w4 >= 0) {
         w_vec <- c(w1, w2, w3, w4)
         err <- get_error(M, M_hat_ls, w_vec)
      } else {
        err = Inf
      }
     
      
      if (err < error_best) {
        wt_best <- w_vec
        error_best <- err
      }
      
    }
  }
}

## best error
print(error_best)

## best weight
print(wt_best)
```



