---
title: "Analysis of Junjian's Data"
author: "Tianyi Liu"
date: "10/05/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir="/Users/gr8lawrence/Documents/Dissertation/conv_opt_code/cleaned_algorithm/PFSM/R")
pacman::p_load(tidyverse, stringr, purrr, furrr, magrittr)
source("./00_core_functions.R")
source("./01_matrix_initiation_functions.R")
source("./02_utility_functions.R")
```

## Seed Control
```{r set.seed}
set.seed(100)
```

## Read Data
We begin with reading in all the bulk, base, and proportion matrices:
```{r read.data, message=FALSE, warning=FALSE}
## directory names 
base_mtx_dir <- "./Real_data/Junjian_01/base_mtx"
bulk_dir <- "./Real_data/Junjian_01/bulk"
proportions_dir <- "./Real_data/Junjian_01/proportions"
output_dir <- "./Real_data/Junjian_01/output"

## all the base matrices
base_ABIS_logFC_a2s <- read_csv(paste(base_mtx_dir, "ABIS_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_immunoStates_logFC_a2s <- read_csv(paste(base_mtx_dir, "immunoStates_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_LM22_logFC_a2s <- read_csv(paste(base_mtx_dir, "LM22_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)
base_tianyi_logFC_a2s <- read_csv(paste(base_mtx_dir, "tianyi_logFC_a2s.csv", sep="/")) %>% rename(gene=X1)

## all the bulk matrices
bulk_markers_sd <- read_tsv(paste(bulk_dir, "bulk_markers_sd.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_markers_sd_q <- read_tsv(paste(bulk_dir, "bulk_markers_sd_q.txt", sep="/")) %>% rename(gene = GeneSymbol) 
bulk_nonmarkers_sd <- read_tsv(paste(bulk_dir, "bulk_nonmkrs_sd.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_nonmarkers_sd_q <- read_tsv(paste(bulk_dir, "bulk_nonmkrs_sd_q.txt", sep="/")) %>% rename(gene = GeneSymbol)
bulk_Z0201 <- read_tsv(paste(bulk_dir, "Z0201_bulkCIBERSORT_a2s.txt", sep="/")) %>% rename(gene = GeneSymbol)

## all the proportion matrices
prop_tianyi <- read_csv(paste(proportions_dir, "CIBERSORTx_Job144_Results_tianyi.csv", sep="/")) 
prop_lm22 <- read_csv(paste(proportions_dir, "CIBERSORTx_Job145_Results_lm22.csv", sep="/"))
prop_immunoStates <- read_csv(paste(proportions_dir, "CIBERSORTx_Job146_Results_immuno.csv", sep="/"))
prop_abis <- read_csv(paste(proportions_dir, "CIBERSORTx_Job147_Results_abis.csv", sep="/"))
```

## Constants
We identify some constants from the data that are applciable to each of the matrices:
```{r get.constants}
n_cell_types <- 6
n_subjects <- 51
```

## Data Pre-processing
Next, we write a series of helper functions to obtain matrices for deconvolution:
```{r helper.function}
get_bulk <- function(bulk_df) {
  # subset the expression and get gene and subject names
  gene_names = bulk_df$gene
  bulk_exp_df = bulk_df %>% 
    select(2:(1 + n_subjects)) 
  subject_names = colnames(bulk_exp_df) %>% 
    substr(2, 5)
  
  # assemble the matrix
  mat = data.matrix(bulk_exp_df)
  rownames(mat) = gene_names
  colnames(mat) = subject_names
  return(mat)
}

#get_bulk(bulk_markers_sd)

get_base <- function(base_df) {
  gene_names = base_df$gene
  base_exp_df = base_df %>% 
    select(2:(1 + n_cell_types))
  cell_type_names = colnames(base_exp_df)
  
  mat = data.matrix(base_exp_df)
  rownames(mat) = gene_names
  colnames(mat) = cell_type_names
  return(mat)
}

#get_base(base_ABIS_logFC_a2s)

get_prop <- function(prop_df) {
  subject_names = prop_df$Mixture
  prop_real_df = prop_df %>% 
    select(2:(1 + n_cell_types))
  cell_type_names = colnames(prop_real_df)
  
  mat = data.matrix(prop_real_df) %>% t
  rownames(mat) = cell_type_names
  colnames(mat) = subject_names
  return(mat)
}

#get_prop(prop_abis)
```

With the new functions we get all the 5 bulk matrices prepared:
```{r bulk.matrices}
bulk_markers_mat <- get_bulk(bulk_markers_sd)
bulk_markers_q_mat <- get_bulk(bulk_markers_sd_q)
bulk_nonmkrs_mat <- get_bulk(bulk_nonmarkers_sd)
bulk_nonmkrs_q_mat <- get_bulk(bulk_nonmarkers_sd_q)
bulk_Z0201_mat <- get_bulk(bulk_Z0201)
```

We also define some tuning parameters:
```{r tuning.vecs}
v_alpha = 2^seq(4, 8, 1)
v_xi = 2^-6
v_beta = 2^seq(2, 10, 2)
```

## Main Helper Functions
We will use the following function to run the algorithm with 5 initial values given a set of tuning parameters and find the best run by the average column-wise correlations between the calculated and the truth:
```{r main.helper.function}
## helper function to run the algorithm given a set of tuning parameter values and a pair of base and bulk matrices
## choose the best run as results
run_PSMF <- function(bulk, base, prop, alpha, xi, beta, k=5) {
  
  base = base_lm22_mat
  bulk = bulk_markers_mat
  prop = get_prop(prop_lm22)
  
  ## best avg cor
  best_cor = 0
  
  ## best solution
  best_sol = NULL
  
  ## get the gene names of base and bulk matrices
  x = rownames(base)
  y = rownames(bulk)
  
  ## genes that are in base but not in bulk matrix
  base_x = setdiff(x, y)
  
  ## genes that are in bulk but not base matrix
  bulk_y = setdiff(y, x)
  
  ## find the coordinates of this gene
  base_x_coords = which(x %in% base_x)
  bulk_y_coords = which(y %in% bulk_y)
  
  ## remove those genes from those coordinates
  G0_upper = base[-base_x_coords, ]
  M = bulk[-bulk_y_coords, ]
  
  ## get the number of unknown genes
  d_len = nrow(M) - nrow(G0_upper)
  
  ## fill G0
  mean_G = mean(base)
  G0_lower = matrix(0L, d_len, n_cell_types)
  G0 = rbind(G0_upper, G0_lower)
  
  ## other parameters
  n_genes = nrow(M)
  n_known_genes = nrow(G0_upper)
  n_good_cell_types = n_cell_types
  
  ## run the algorithm with k different initial values and try to find the one with the best correlation
  for (i in 1:k) {
    G_init_lower = matrix(rchisq(d_len * n_cell_types, mean_G),
                        d_len,
                        n_cell_types)
  
    G_init = rbind(G0_upper, G_init_lower)
    C_init = get_C(n_cell_types, n_subjects)
    
    solve_tmp = PSMF_solve(
      M = M,
      G_0 = G0,
      G_init = G_init,
      C_init = C_init,
      n_markers = n_known_genes,
      n_good_cell_types = n_good_cell_types,
      alpha = alpha,
      xi = xi,
      beta = beta
    )
    
    C_hat_tmp = solve_tmp$C_hat
    
    ## calculate the average correlation per cell type
    # cor_tmp = colwise_cor(t(C_hat_tmp), t(prop))
    
    ## calculate the average correlation per sample
    cor_tmp = colwise_cor(C_hat_tmp, prop)
    
    if (is.na(cor_tmp)) {
      print("Unable to calculate correlations due to no variation in sample/cell type proportion estimates.")
      cor_tmp = 0
    }
    
    if (cor_tmp > best_cor) {
      best_cor = cor_tmp
      best_sol = solve_tmp
    }
    
  }
  
  if (best_cor == 0) {
    print("None of the solutions has positive correlation with the truth!")
    res = list(cor = 0, sol = NULL)
    return(res)
  } else {
    #print(paste0("Best correlation = ", round(best_cor, 3)))
    res = list(cor = best_cor, sol = best_sol)
    return(res)
  }

}
```

Then, we write another upper-level function to control the input of tuning parameters:
```{r upper.function}
tune_PSMF <- function(v_alpha, v_xi, v_beta, ...) {
  
  ## the best correlation
  best_cor = 0
  
  ## the best solution
  best_sol = NULL
  
  for (i in 1:length(v_alpha)) {
    for (j in 1:length(v_xi)) {
      for (k in 1:length(v_beta)) {
        
        ## get the tuning parameters
        alpha = v_alpha[i]
        xi = v_xi[j]
        beta = v_beta[k]
        
        ## run the algorithm with the above parameters
        solve = run_PSMF(..., alpha=alpha, xi=xi, beta=beta)
        
        ## collect the results
        cor_tmp = solve$cor
        
        ## replace the best result if the result improves
        if (cor_tmp > best_cor) {
          best_cor = cor_tmp
          best_sol = solve
        }
        
      }
    }
  }
  
  if (best_cor == 0) {
    print("Overall, none of the solutions has positive correlation with the truth!")
    res = list(cor = 0, sol = NULL, y_hat = NULL)
    return(res)
  } else {
    print(paste0("Best correlation = ", round(best_cor, 3)))
    res = list(cor = best_cor, sol = best_sol)
    return(res)
  }
  
}
```

## ABIS
We run the deconvolution algorithm using the ABIS matrix as base matrix on 5 different bulk matrices
```{r run.ABIS}
## get the abis base and prop matrix
base_abis_mat <- get_base(base_ABIS_logFC_a2s)
prop_abis_mat <- get_prop(prop_abis)

## run deconvolution on all bulks
ptm <- proc.time()
abis_tune_1 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_mat,
                         base=base_abis_mat,
                         prop=prop_abis_mat)

## save the estimated bulk
abis_y_hat1 <- abis_tune_1$sol$sol$G_hat %*% abis_tune_1$sol$sol$C_hat
write_csv(as_tibble(abis_y_hat1), paste(output_dir, "abis_all_markers.csv", sep="/"))

print("Completed the first bulk matrix.")

abis_tune_2 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_q_mat,
                         base=base_abis_mat,
                         prop=prop_abis_mat)

abis_y_hat2 <- abis_tune_2$sol$sol$G_hat %*% abis_tune_2$sol$sol$C_hat
write_csv(as_tibble(abis_y_hat2), paste(output_dir, "abis_interq_markers.csv", sep="/"))

print("Completed the second bulk matrix.")

abis_tune_3 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_mat,
                         base=base_abis_mat,
                         prop=prop_abis_mat)

abis_y_hat3 <- abis_tune_3$sol$sol$G_hat %*% abis_tune_3$sol$sol$C_hat
write_csv(as_tibble(abis_y_hat3), paste(output_dir, "abis_all_nonmkrs.csv", sep="/"))

print("Completed the third bulk matrix.")

## This is the estimated bulk we wanted
abis_tune_4 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_q_mat,
                         base=base_abis_mat,
                         prop=prop_abis_mat)

abis_y_hat4 <- abis_tune_4$sol$sol$G_hat %*% abis_tune_4$sol$sol$C_hat
write_csv(as_tibble(abis_y_hat4), paste(output_dir, "abis_interq_nonmkrs.csv", sep="/"))

# alpha = v_alpha[1]
# xi = v_xi[1]
# beta = v_beta[1]
# 
# run_1 <- run_PSMF(bulk = bulk_nonmkrs_q_mat, base = base_abis_mat,
#                   prop = prop_abis_mat, alpha = alpha, xi = xi, beta = beta)

print("Completed the fourth bulk matrix.")

abis_tune_5 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_Z0201_mat,
                         base=base_abis_mat,
                         prop=prop_abis_mat)

abis_y_hat5 <- abis_tune_5$sol$sol$G_hat %*% abis_tune_5$sol$sol$C_hat
write_csv(as_tibble(abis_y_hat5), paste(output_dir, "abis_interq_Z0201.csv", sep="/"))

print("Completed the fifth bulk matrix.")

proc.time() - ptm
```

## immonuoStates
```{r run.immunoStates, warning=FALSE}
## get the immunoStates base and prop matrix
base_immuno_mat <- get_base(base_immunoStates_logFC_a2s)
prop_immuno_mat <- get_prop(prop_immunoStates)

## run deconvolution on all bulks
ptm <- proc.time()
immuno_tune_1 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_mat,
                         base=base_immuno_mat,
                         prop=prop_immuno_mat)

## save the estimated bulk
if (immuno_tune_1$cor > 0) {
  immuno_y_hat1 <- immuno_tune_1$sol$sol$G_hat %*% immuno_tune_1$sol$sol$C_hat
  write_csv(as_tibble(immuno_y_hat1), paste(output_dir, "immuno_all_markers.csv", sep="/"))
}


print("Completed the first bulk matrix.")

immuno_tune_2 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_q_mat,
                         base=base_immuno_mat,
                         prop=prop_immuno_mat)

if (immuno_tune_2$cor > 0) {
  immuno_y_hat2 <- immuno_tune_2$sol$sol$G_hat %*% immuno_tune_2$sol$sol$C_hat
  write_csv(as_tibble(immuno_y_hat2), paste(output_dir, "immuno_interq_markers.csv", sep="/"))
}

print("Completed the second bulk matrix.")

immuno_tune_3 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_mat,
                         base=base_immuno_mat,
                         prop=prop_immuno_mat)

if (immuno_tune_3$cor > 0) {
  immuno_y_hat3 <- immuno_tune_3$sol$sol$G_hat %*% immuno_tune_3$sol$sol$C_hat
  write_csv(as_tibble(immuno_y_hat3), paste(output_dir, "immuno_all_nonmkrs.csv", sep="/"))
}

print("Completed the third bulk matrix.")

## This is the estimated bulk we wanted
immuno_tune_4 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_q_mat,
                         base=base_immuno_mat,
                         prop=prop_immuno_mat)

if (immuno_tune_4$cor > 0) {
  immuno_y_hat4 <- immuno_tune_4$sol$sol$G_hat %*% immuno_tune_4$sol$sol$C_hat
  write_csv(as_tibble(immuno_y_hat4), paste(output_dir, "immuno_interq_nonmkrs.csv", sep="/"))
}

print("Completed the fourth bulk matrix.")

immuno_tune_5 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_Z0201_mat,
                         base=base_immuno_mat,
                         prop=prop_immuno_mat)

if (immuno_tune_5$cor > 0) {
  immuno_y_hat5 <- immuno_tune_5$sol$sol$G_hat %*% immuno_tune_5$sol$sol$C_hat
  write_csv(as_tibble(immuno_y_hat5), paste(output_dir, "immuno_interq_Z0201.csv", sep="/"))
}

print("Completed the fifth bulk matrix.")

proc.time() - ptm
```

## LM22
```{r run.LM22, warning=FALSE}
## get the abis base and prop matrix
base_lm22_mat <- get_base(base_LM22_logFC_a2s)
prop_lm22_mat <- get_prop(prop_lm22)

## run deconvolution on all bulks
ptm <- proc.time()
lm22_tune_1 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_mat,
                         base=base_lm22_mat,
                         prop=prop_lm22_mat)

## save the estimated bulk
if (lm22_tune_1$cor > 0) {
  lm22_y_hat1 <- lm22_tune_1$sol$sol$G_hat %*% lm22_tune_1$sol$sol$C_hat
  write_csv(as_tibble(lm22_y_hat1), paste(output_dir, "lm22_all_markers.csv", sep="/"))
}

# alpha = v_alpha[1]
# xi = v_xi[1]
# beta = v_beta[1]
# 
# run_1 <- run_PSMF(bulk = bulk_nonmkrs_q_mat, base = base_lm22_mat,
#                   prop = prop_lm22_mat, alpha = alpha, xi = xi, beta = beta)


print("Completed the first bulk matrix.")

lm22_tune_2 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_q_mat,
                         base=base_lm22_mat,
                         prop=prop_lm22_mat)

if (lm22_tune_2$cor > 0) {
  lm22_y_hat2 <- lm22_tune_2$sol$sol$G_hat %*% lm22_tune_2$sol$sol$C_hat
  write_csv(as_tibble(lm22_y_hat2), paste(output_dir, "lm22_interq_markers.csv", sep="/"))
}

print("Completed the second bulk matrix.")

lm22_tune_3 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_mat,
                         base=base_lm22_mat,
                         prop=prop_lm22_mat)

if (lm22_tune_3$cor > 0) {
  lm22_y_hat3 <- lm22_tune_3$sol$sol$G_hat %*% lm22_tune_3$sol$sol$C_hat
  write_csv(as_tibble(lm22_y_hat3), paste(output_dir, "lm22_all_nonmkrs.csv", sep="/"))
}


print("Completed the third bulk matrix.")

## This is the estimated bulk we wanted
lm22_tune_4 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_q_mat,
                         base=base_lm22_mat,
                         prop=prop_lm22_mat)

if (lm22_tune_4$cor > 0) {
  lm22_y_hat4 <- lm22_tune_4$sol$sol$G_hat %*% lm22_tune_4$sol$sol$C_hat
  write_csv(as_tibble(lm22_y_hat4), paste(output_dir, "lm22_interq_nonmkrs.csv", sep="/"))
}

print("Completed the fourth bulk matrix.")

lm22_tune_5 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_Z0201_mat,
                         base=base_lm22_mat,
                         prop=prop_lm22_mat)

if (lm22_tune_5$cor > 0) {
  lm22_y_hat5 <- lm22_tune_5$sol$sol$G_hat %*% lm22_tune_5$sol$sol$C_hat
  write_csv(as_tibble(lm22_y_hat5), paste(output_dir, "lm22_interq_Z0201.csv", sep="/"))
}


print("Completed the fifth bulk matrix.")

proc.time() - ptm
```

## tianyi
```{r run.tianyi, warning=FALSE}
## get the abis base and prop matrix
base_tianyi_mat <- get_base(base_tianyi_logFC_a2s)
prop_tianyi_mat <- get_prop(prop_tianyi)

## get the genes in the base matrix
gene_tianyi <- rownames(base_tianyi_mat)

## run deconvolution on all bulks
ptm <- proc.time()
tianyi_tune_1 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_mat,
                         base=base_tianyi_mat,
                         prop=prop_tianyi_mat)

## save the estimated bulk
if (tianyi_tune_1$cor > 0) {
  tianyi_y_hat1 <- tianyi_tune_1$sol$sol$G_hat %*% tianyi_tune_1$sol$sol$C_hat
  write_csv(as_tibble(tianyi_y_hat1), paste(output_dir, "tianyi_all_markers.csv", sep="/"))
}

print("Completed the first bulk matrix.")

tianyi_tune_2 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_markers_q_mat,
                         base=base_tianyi_mat,
                         prop=prop_tianyi_mat)

if (tianyi_tune_2$cor > 0) {
  tianyi_y_hat2 <- tianyi_tune_2$sol$sol$G_hat %*% tianyi_tune_2$sol$sol$C_hat
  write_csv(as_tibble(tianyi_y_hat2), paste(output_dir, "tianyi_interq_markers.csv", sep="/"))
}

print("Completed the second bulk matrix.")

tianyi_tune_3 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_mat,
                         base=base_tianyi_mat,
                         prop=prop_tianyi_mat)

if (tianyi_tune_3$cor > 0) {
  tianyi_y_hat3 <- tianyi_tune_3$sol$sol$G_hat %*% tianyi_tune_3$sol$sol$C_hat
  write_csv(as_tibble(tianyi_y_hat3), paste(output_dir, "tianyi_all_nonmkrs.csv", sep="/"))
}

print("Completed the third bulk matrix.")

## This is the estimated bulk we wanted
tianyi_tune_4 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_nonmkrs_q_mat,
                         base=base_tianyi_mat,
                         prop=prop_tianyi_mat)

if (tianyi_tune_4$cor > 0) {
  tianyi_y_hat4 <- tianyi_tune_4$sol$sol$G_hat %*% tianyi_tune_4$sol$sol$C_hat
  write_csv(as_tibble(tianyi_y_hat4), paste(output_dir, "tianyi_interq_nonmkrs.csv", sep="/"))
}

print("Completed the fourth bulk matrix.")

tianyi_tune_5 <- tune_PSMF(v_alpha=v_alpha,
                         v_xi=v_xi,
                         v_beta=v_beta,
                         bulk=bulk_Z0201_mat,
                         base=base_tianyi_mat,
                         prop=prop_tianyi_mat)

if (tianyi_tune_5$cor > 0) {
  tianyi_y_hat5 <- tianyi_tune_5$sol$sol$G_hat %*% tianyi_tune_5$sol$sol$C_hat
  write_csv(as_tibble(tianyi_y_hat5), paste(output_dir, "tianyi_interq_Z0201.csv", sep="/"))
}

print("Completed the fifth bulk matrix.")

proc.time() - ptm
```

