---
title: "Comparison between NITUMID and our algorithm"
author: "Tianyi Liu"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/gr8lawrence/Documents/Dissertation/conv_opt_code/cleaned_algorithm/PFSM/R")
pacman::p_load(tidyverse, stringr, purrr, furrr, magrittr, NITUMID)
source("./core_functions.R")
source("./matrix_initiation_functions.R")
source("./utility_functions.R")
```

## NITUMID

NITUMID is a software that uses similar update functions for matrix decomposition. We will test our code by comparing the step-wise updates between our and NITUMID's result.

In NITUMID, `Update_W_Tang` is used to update $\bf{G}$ and `Update_H_Tang` $\bf{C}$.

We need to keep the dimensions small for quick calculation. Using 
```{r}
m = 12 # total number of genes
m_kn = 0 # total number of known genes
n = 1 # sample size
p = 5 # number of CTs
p_good = 0 # non-missing CTs
```

First, we generate some small batch of G and C:
```{r}
set.seed(100)
#G_itr = get_G(m, n, p, m_kn, p_good) # get an initial G
G_itr = matrix(abs(rnorm(m * p, 0, 1)), m, p)
C_itr = get_C(p, n) # get an initial C
E = matrix(runif(m * n), m, n)
M = G_itr %*% C_itr + E
A = matrix(1L, m, p)
G0 = matrix(0L, m, p)
G0[1:m_kn, 1:p_good] = G_itr[1:m_kn, 1:p_good]
```

Then, we need to let $\alpha = 0$ for our algorithm and $\lambda = 0$ for NTUMID.

```{r}
G_new_Tang = Update_W_Tang(
  Y = M,
  W_itr = G_itr,
  H_itr = C_itr,
  beta_W = 4,
  lam_W = 0,
  A = A
)

G_new_ours = update_G_TL(
  G_old = G_itr,
  M = M,
  C_new = C_itr,
  G_0 = G0,
  n_markers = m_kn,
  n_good_cell_types = p_good,
  alpha = 0,
  xi = 4
)


all.equal(G_new_Tang, G_new_ours)

C_new_Tang = Update_H_Tang(
  Y = M,
  W_itr = G_itr,
  H_itr = C_itr,
  eta = 10
)

C_new_ours = update_C_TL(
  C_old = C_itr,
  M = M,
  G_old = G_itr,
  beta = 10
)

all.equal(C_new_Tang, C_new_ours)
```
