---
title: "Simulation Round 1"
author: "Tianyi Liu"
date: "10/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/Users/gr8lawrence/Documents/Dissertation/conv_opt_code/cleaned_algorithm/PFSM/R/simulation_results")
pacman::p_load(tidyverse)
```

## Varying Sample Size
Number of genes = 20; ratio of marker genes = 0.5
### Read in data
```{r read.data}
load("sim_round_4.RData")
res <- as_tibble(res) 
colnames(res) <- c("n", "m", "p", "alpha", "xi", "beta", "run", "cor", "mse")
res
```

### Find each run's best result for each combination of "n", "m", and "p"

The initial values are chosen not so close to the true value:
```{r vary.samp.size.cor}
ggplot(res, aes(x = as.factor(n), y = cor)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(desc(cor)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(n), y = cor)) +
  geom_boxplot()
```

Also, plot the distribution of MSEs:
```{r vary.samp.size.mse}
ggplot(res, aes(x = as.factor(n), y = mse)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(mse) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(n), y = mse)) +
  geom_boxplot()
```

## Varying Number of Genes
Sample size = 20; ratio of marker genes = 0.5
### Read in data
```{r read.data2}
load("sim_round_5.RData")
res <- as_tibble(res) 
colnames(res) <- c("n", "m", "p", "alpha", "xi", "beta", "run", "cor", "mse")
res
```

### Find each run's best result for each combination of "n", "m", and "p"

The initial values are chosen not so close to the true value:
```{r vary.num.gene.cor}
ggplot(res, aes(x = as.factor(m), y = cor)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(desc(cor)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(m), y = cor)) +
  geom_boxplot()
```

Also, plot the distribution of MSEs:
```{r vary.num.gene.mse}
ggplot(res, aes(x = as.factor(m), y = mse)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(mse) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(m), y = mse)) +
  geom_boxplot()
```

## Varying Ratio of Marker Genes
Number of samples = number of genes = 50
### Read in data
```{r read.data3}
load("sim_round_6.RData")
res <- as_tibble(res) 
colnames(res) <- c("n", "m", "p", "alpha", "xi", "beta", "run", "cor", "mse")
res
```

The initial values are chosen not so close to the true value:
```{r vary.ratio.cor}
ggplot(res, aes(x = as.factor(m), y = cor)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(desc(cor)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(m), y = cor)) +
  geom_boxplot()
```

Also, plot the distribution of MSEs:
```{r vary.ratio.mse}
ggplot(res, aes(x = as.factor(m), y = mse)) +
  geom_boxplot() +
  facet_grid(alpha ~ beta)

best_res <- res %>% 
  group_by(n, m, p, run) %>% 
  arrange(mse) %>% 
  filter(row_number() == 1) %>% 
  ungroup() 

best_res

ggplot(best_res, aes(x = as.factor(m), y = mse)) +
  geom_boxplot()
```